
```{r, echo=FALSE}
#set global opts
opts_chunk$set(message=FALSE,warning=FALSE,echo=FALSE)
```

```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE,results='asis'}
=======
opts_chunk$set(message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE,fig.height=8,fig.width=11,fig.res=300)
```

**What does the data look like?**

```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE,results='asis',}
################################################################
#Step 1 - Set up data environment and load in data
################################################################

#load in packages
require(ggplot2)
require(chron)
require(reshape)

#set gitpath
gitpath<-"C:/Users/Jorge/Documents/Selectivity/"

#source functions
source(paste(gitpath,"functions.R",sep=""))

#Set working directory
droppath<-"C:/Users/Jorge/Dropbox/"
setwd(droppath)

##Read in data
dat<-read.csv(paste(droppath,"Thesis//Maquipucuna_SantaLucia/Data2013/csv/CompetitionFeeders.csv",sep=""))

#Make all entries capital
dat$Sex<-toupper(dat$Sex)

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv("Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",row.names=1)

dat<-dat[,1:12]
require(xtable)
```


```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE,results='asis'}
#####################################
#Step 2 - Data Cleaning and Sampling
#####################################

#Which dates need to be matched?
vid_totals_date<-aggregate(dat$Video,list(dat$Elevation,dat$Treatment,dat$Date,dat$Replicate),function(x) nlevels(droplevels(x)))
vid_totals_date<-cast(vid_totals_date,Group.1 + Group.3 + Group.4~Group.2)

colnames(vid_totals_date)<-c("Elevation","Date","Replicate(O_R)","High Feeder","Low Feeder")
require(xtable)
```


```{r}
require(chron)
dat<-dat[!dat$Species %in% "UKWN",]

#Create time columns
dat$Time.End<-times(dat$Time.End)
dat$Time.Begin<-times(dat$Time.Begin)

#Find time difference 
dat$Time_Feeder_Obs<-dat$Time.End - dat$Time.Begin

#Get any rownumbers that are negative, these need to be fixed. 
#dat[which(dat$Time_Feeder_Obs < 0),]
```


```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE}
#average visits per hour
S_H<-table(hours(dat$Time.Begin),dat$Species)

#Create a species list for each video

####Match each trial together, trials are done on the same day at the same elevation
#Split data into a list, with each compenent being one trial pair
Trials<-split(dat, list(dat$Elevation,dat$Date,dat$Replicate),drop=TRUE)

#How many hours of video

#####Just for data clarity remove any trials that down have high and low value data entered
#Get number of levels per trial
levels.trial<-lapply(Trials,function(x) nlevels(factor(x$Treatment)))

#Only use trials that have a high and low, ie levels=2
complete.trial<- Trials[levels.trial ==2]
```


```{r,message=FALSE,warning=FALSE,cache=TRUE,results='asis',echo=FALSE}
####Within trial metrics per species

Tdata<-lapply(complete.trial,function(x){
  a<-selective(x)
  b<-bph(x)
  d<-avgF(x)
  dat.trials<-merge(merge(a,b),d)
  Elevation=unique(x$Elevation)
  Date=unique(x$Date)
  Replicate=unique(x$Replicate)
  #Richness at that feeder
  Trichness<-length(levels(droplevels(x$Species)))
  #Total visits
  Tvisits<-nrow(x)
  out<-data.frame(dat.trials,Elevation,Date,Replicate,Richness=Trichness,Tvisits)
  return(out)})

#Bind dataset to a dataframe
selective.matrix<-rbind.fill(Tdata)

selective.matrix$Time_High<-times(selective.matrix$Time_High)
selective.matrix$Time_Low<-times(selective.matrix$Time_Low)
selective.matrix$Total_Time<-selective.matrix$Time_High + selective.matrix$Time_Low

#add total minutes feeding as a weight
selective.matrix$Minutes_High<-minutes(selective.matrix$Time_High) + seconds(selective.matrix$Time_High)/60
selective.matrix$Minutes_Low<-minutes(selective.matrix$Time_Low) + seconds(selective.matrix$Time_Low)/60
selective.matrix$Minutes_Total<-selective.matrix$Minutes_Low+selective.matrix$Minutes_High

#Add month column
selective.matrix$MonthA<-format(as.POSIXct(selective.matrix$Date,format="%m/%d/%Y"),"%b")

#I think i the weighted regression is weird, just take out data where the bird fed less than 

# plot(ecdf(selective.matrix$Minutes_Total))
# 
# ecdf(selective.matrix$Minutes_Total) (1)
# 
# hist(selective.matrix$Minutes_Total,breaks=seq(0,40,.5))
# hist(selective.matrix$Minutes_High,breaks=seq(0,40,.5))
# hist(selective.matrix$Minutes_Low,breaks=seq(0,40,.5))

#Take out birds feeding less than 1min over the 6hours
selective.matrix<-selective.matrix[selective.matrix$Minutes_Total > 1,]
print(xtable(head(selective.matrix)),type="html")

```

***********


```{r,message=FALSE,warning=FALSE,cache=TRUE}
<<<<<<< HEAD
=======
#Birds should prefer high value resources
>>>>>>> b11be5f4da282c9a3a1b948f5c98b38c1ca1cd11

#Optimal foraging says that individuals should occupy patches at a rate equal to their quality
sH<-sum(selective.matrix$Minutes_High)
sL<-sum(selective.matrix$Minutes_Low)

#Optimal foraging says its should be three
<<<<<<< HEAD
=======

>>>>>>> b11be5f4da282c9a3a1b948f5c98b38c1ca1cd11

#Competition keeps birds from occupying higher quality patch

#Effect of increasing richness on selectivity
ggplot(selective.matrix,aes(x=Richness,y=Selectivity)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total)) + scale_x_continuous(breaks=seq(0,9,1))

#By species
ggplot(selective.matrix,aes(x=Richness,y=Selectivity)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total)) + scale_x_continuous(breaks=seq(0,9,1)) + facet_wrap(~Species)

<<<<<<< HEAD
#Effects of increasing visits on selectivity
ggplot(selective.matrix,aes(x=Tvisits,y=Time_High)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total))

#Effect of increasing time on the high value resource on number of other species.
ggplot(selective.matrix,aes(x=as.numeric(Richness-1),y=Minutes_High)) + geom_point() + stat_smooth(method="lm") + scale_x_continuous(breaks=seq(0,8,1)) + facet_wrap(~Species,scales="free")
```

If competition was not a factor influecing hummingbird feeding selectivity, we would expect that birds should feed three times as long on the high value resource as the low. 
=======
# #Effects of increasing visits on selectivity
# ggplot(selective.matrix,aes(x=Tvisits,y=Selectivity)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total))
```

#Correlations among data
------------------------

```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE}
#pairs plot
require(GGally)
ggpairs(selective.matrix[,c("Selectivity","bph","avgF","Elevation","Minutes_High","Minutes_Low","Minutes_Total")])

##########################
#merge with morphology
##########################
#PCA of trait space

# Hum Standardized variables, what to do about NA's?
#Standard the matrix to correct for different units by subtracting the means and dividing by sd
zscore <- apply(hum.morph[,c("Bill","Mass","WingChord","Tarsus_Length","Nail_Length","Wing_Loading")], 2, function(x){
  y<-(x - mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)
  return(y)
})
rownames(zscore)<-hum.morph$English

#Need to figure out what to do about Na's, we could use closely related species?
trait_pc<-prcomp(na.omit(zscore))

#View Biplot of PC Space
biplot(trait_pc)

#bind loadings 1 and 2 to dataframe
hum_load<-trait_pc$x[,c("PC1","PC2")]
rownames(hum_load)<-rownames(zscore)

hum.morph<-merge(hum.morph,hum_load,by.x="English",by.y="row.names")

#merge PCA and original data with selectivity
selective.matrix<-merge(selective.matrix,hum.morph[,!colnames(hum.morph) %in% "Species"],by.x="Species",by.y="English")

#Weighted average of selectivity
ws<-sapply(split(selective.matrix,selective.matrix$Species),function(x){
  weighted.mean(x$Selectivity,x$Total_Time)
})

selective.matrix<-merge(selective.matrix,data.frame(weighted.selectivity=ws),by.x="Species",by.y="row.names")

#average metrics across species
avgStat<-aggregate(selective.matrix[,c("avgF","bph")],by=list(selective.matrix$Species),mean,na.rm=TRUE)
colnames(avgStat)<-c("Species","avgF")

#merge with weighted selectivity
avgStat<-merge(ws,avgStat,by.x="row.names",by.y="Species")
colnames(avgStat)<-c("Species","W.Selectivity","avgF","bph")
rownames(avgStat)<-avgStat$Species

pcaStat<-prcomp(avgStat[,-1],scale=TRUE)
biplot(pcaStat)
```

Hypothesis 1:
------------
Species should be most selective at the center of their range. Following the abundance center hypothesis, we expect the highest quality territory at the center of species elevation range. Individuals in non-suitable habitat, or weakly suitable habitat, may be inferior competitions due to mal-adaptation to the environment.

```{r,message=FALSE,warning=FALSE,cache=TRUE}
#get only species that have atleast 2min
sumT<-aggregate(selective.matrix$Minutes_Total,list(selective.matrix$Species),sum)
speciesSkip<-sumT[sumT$x < 2,]$Group.1

#unweighted
p<-ggplot(selective.matrix[!selective.matrix$Species %in% speciesSkip, ],aes(x=as.numeric(Elevation),Selectivity,col=Species)) + geom_point(size=3) + facet_wrap(~Species) + stat_smooth(method="glm",aes(group=1),family="binomial")
ggsave(paste(gitpath,"Figures//Selectivity_Elevation_Unweighted.svg",sep=""),height=8,width=15)

#weighted
p<-ggplot(selective.matrix[!selective.matrix$Species %in% speciesSkip & selective.matrix$Minutes_Total > 1,],aes(x=as.numeric(Elevation),Selectivity,col=Species,size=Minutes_Total)) + geom_point() + facet_wrap(~Species)
p  + stat_smooth(method="glm",family="binomial") + theme_bw() + xlab("Elevation") + scale_x_continuous(breaks=as.numeric(levels(factor(dat$Elevation))))
ggsave(paste(gitpath,"Figures//Selectivity_Elevation_Unweighted.svg",sep=""),height=8,width=15)

# # #weighted and time
# p<-ggplot(selective.matrix[!selective.matrix$Species %in% speciesSkip,],aes(x=as.numeric(Elevation),Selectivity,col=MonthA,size=Minutes_Total)) + geom_point() + facet_wrap(~Species)
# p
# p  + geom_smooth(method="glm",family="binomial",aes(weight=Minutes_Total)) + theme_bw() + xlab("Elevation") + stat_smooth()

## Write selectivity tables to file
write.csv(selective.matrix,paste(droppath,"Thesis//Maquipucuna_SantaLucia/Results/Selectivity/Selectivity_Elevation.csv",sep=""))
```

```{r,message=FALSE,warning=FALSE,cache=TRUE}
#######################################
#Selectivity, Phylogeny and Morphology
#######################################

#Selectivity Descriptors for each species
ggplot(selective.matrix[!selective.matrix$Species %in% speciesSkip,],aes(x=Species,Selectivity)) + geom_boxplot() + theme(axis.text.x=element_text(angle=-90,vjust=-.1))

#aggregate
wss<-aggregate(selective.matrix$weighted.selectivity,by=list(selective.matrix$Species,selective.matrix$PC1,selective.matrix$PC2),mean)
colnames(wss)<-c("Species","PC1","PC2","weighted.selectivity")

biplot(trait_pc)

ggplot(wss,aes(x=PC2,y=weighted.selectivity)) + geom_point() + geom_smooth(method="glm",family="binomial") + geom_text(aes(label=Species),size=2)
ggplot(wss,aes(x=PC1,y=weighted.selectivity)) + geom_point() + geom_smooth(method="glm",family="binomial") + geom_text(aes(label=Species),size=2)
# 
# p<-ggplot(selective.matrix,aes(x=PC1,y=Selectivity,size=Minutes_Total,label=Species)) + geom_point() + stat_smooth(method="glm",link="binomial",aes(weight=Minutes_Total))
# p
# 
# p<-ggplot(selective.matrix,aes(x=PC2,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",link="binomial",aes(weight=Minutes_Total,group=1))
# p

#########################################################
#Selectivity as function of distance to nearest range edge
##########################################################

#Bring in transect data, see Maquipucuna Repo
transect<-read.csv(paste(droppath,"Thesis//Maquipucuna_SantaLucia//Results//HummingbirdTransects//HumTransectRows.csv",sep=""),row.names=1)

#just get summer transect data
head(transect)

levels(transect$Hummingbird.Species)[levels(transect$Hummingbird.Species) %in% c("violet-tailed Slyph","VIolet-tailed Slyph","Violet-tailed Slyph")]<-"Violet-tailed Sylph"
levels(transect$Hummingbird.Species)[levels(transect$Hummingbird.Species) %in% c("Booted Racketail")]<-"Booted Racket-tail"

transect<-transect[transect$Month %in% c(6,7,8),]
dim(transect)

#Range extent for each species
ggplot(transect,aes(x=ele,fill=Hummingbird.Species)) + geom_histogram() + facet_wrap(~Hummingbird.Species)

rangeLim<-aggregate(transect$ele,list(transect$Hummingbird.Species),range,na.rm=TRUE)
rangeLim<-data.frame(rangeLim$Group.1,rangeLim$x)
colnames(rangeLim)<-c("Species","Min","Max")

ggplot(rangeLim,aes(x=Species,ymin=Min,ymax=Max)) + geom_linerange() + coord_flip()

distU<-apply(selective.matrix,1,function(x){
  
  #If species doesn't exist in transect, skip
  if(sum(rangeLim$Species %in% x["Species"])==0) {
    r<-NA
    names(r)<-NA
    return(r)
  }
  
  maxR<-abs(as.numeric(x["Elevation"]) - as.numeric(rangeLim[rangeLim$Species %in% x["Species"],]$Max))
  minR<-abs(as.numeric(x["Elevation"]) - as.numeric(rangeLim[rangeLim$Species %in% x["Species"],]$Min))

    
  if(maxR > minR) {return(data.frame(UP="Dist_Upper",RDist=maxR))}
  if(minR > maxR) {return(data.frame(UP="Dist_Lower",RDist=minR))}
  
})

#append to dataset
selective.matrix<-data.frame(selective.matrix,rbind.fill(distU))

#Selectivity and distance to range edge.
ggplot(selective.matrix,aes(RDist,Selectivity,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1)) 

##########################################
#Selectivity as a function of body size
##########################################

p<-ggplot(selective.matrix,aes(x=Mass,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1))
p

##########################################
#Selectivity as a function of difference in body size
##########################################

#Species list for each trial.

sp.lists<-sapply(complete.trial,function(x){
  levels(droplevels(x$Species))
})

#get body size lists for each trial
mass.lists<-lapply(sp.lists,function(x){
  hum.morph[hum.morph$English %in% x, "Mass"]
})

#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$MassD<-sapply(1:nrow(selective.matrix),function(y){
  
  x<-selective.matrix[y,]
  #find the index in the list
  index<-paste(paste(x["Elevation"],x[["Date"]],sep="."),x[["Replicate"]],sep=".")
  
  mass_T<-mass.lists[names(mass.lists) %in% index]
  
  #mass of the species minus max
  massD<-x[["Mass"]] - max(unlist(mass_T))
  return(massD)})

p<-ggplot(selective.matrix,aes(x=MassD,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1))

#####################################################
#Compare Selectivity to Available Resource
#####################################################

#read in flower totals from FlowerTransects.R
fltransects<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerTransects/CleanedHolgerTransect.csv",sep=""),row.names=1)

fl<-aggregate(fltransects$Total_Flowers,by=list(fltransects$month,fltransects$Elevation.Begin,fltransects$Elevation.End),sum,na.rm=TRUE)

colnames(fl)<-c("Month","Elevation.Begin","Elevation.End","TotalFlowers")

selective.matrix$Resources<-sapply(1:nrow(selective.matrix),function(y){
  
  #get row
  x<-selective.matrix[y,]
  
  #What month
  m<-as.numeric(months(as.character(x$Date)))
  
  #Total flowers at that elevation
  out<-fl[(fl$Elevation.Begin %in% x$Elevation|fl$Elevation.End %in% x$Elevation) & fl$Month %in% m ,]
  
  #Average of the transects above and below
  return(mean(out$TotalFlowers))
  
  })

p<-ggplot(selective.matrix,aes(x=Resources,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1))
p + facet_wrap(~Species)


# ############Quick look at temperature and elevation
# require(scales)
# ggplot(dat,aes(x=chron(Time.Begin),y=Temp,col=factor(Elevation))) + geom_smooth(se=FALSE) + scale_x_chron(format="%H")

#################################
#Species Presence and Time
##################################

#Create overall date stamp
dat$Time_Stamp<-as.POSIXct(chron(dates=as.character(dat$Date),dat$Time.Begin))

#Time and species occurence, facetted by elevation
ggplot(dat,aes(x=strptime(dat$Time.Begin,"%H:%M"),fill=Species)) + geom_histogram(position="dodge") + facet_wrap(~Elevation)
ggsave("Thesis//Maquipucuna_SantaLucia/Results/TimeofDayElevation.svg",height=11,width=8,dpi=300)

#Overall Month_Day and Elevation
ggplot(dat,aes(y=factor(Elevation),x=dat$Time_Stamp,col=Species)) + geom_point(size=3) + scale_x_datetime() + facet_wrap(~Species)
ggsave("Thesis//Maquipucuna_SantaLucia/Results/DateElevation.svg",height=11,width=8,dpi=300)

```

Supplamentary Figures
=====================

```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE}

#Species richness and identity at each elevation
sp_matrixHL<-(table(dat$Species,dat$Elevation,dat$Treatment) >= 1) * 1

#View species at each elevation and treatment
m.sp_m<-melt(sp_matrixHL)
colnames(m.sp_m)<-c("Species","Elevation","Treatment","Presence")

#turn 0's to NA's just for plotting
m.sp_m[m.sp_m$Presence==0,"Presence"]<-NA

#richness across feeders
p<-ggplot(m.sp_m,aes(y=Species,x=factor(Elevation),fill=as.factor(Presence)))+ geom_tile() + theme_bw() + scale_fill_discrete(na.value="white")
p + labs(fill="Present",x="Elevation")
ggsave(paste(gitpath,"Figures/RangeExtentFeeders.svg",sep=""),dpi=300,height=8,width=11)
```


```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE}
#Total Time per species
Total_Time_Species<-aggregate(dat$Time_Feeder_Obs,by=list(dat$Species),sum,na.rm=TRUE) 
colnames(Total_Time_Species)<-c("Species","TotalTime")
Total_Time_Species$Time<-minutes(Total_Time_Species$TotalTime)+seconds(Total_Time_Species$TotalTime)/60

ggplot(Total_Time_Species,aes(Species,Time)) + geom_bar() + theme_bw() + ylab("Minutes on Feeders") + theme(axis.text.x=element_text(angle=-90,vjust=-.1))
```

Figure 3. Average duration of feeding event for each species

```{r,message=FALSE,warning=FALSE,cache=TRUE,echo=FALSE}
#mean time feeding bout
Mean_Time_Species<-aggregate(dat$Time_Feeder_Obs,by=list(dat$Species),mean,na.rm=TRUE) 
colnames(Mean_Time_Species)<-c("Species","Mean_Time")
ggplot(Mean_Time_Species,aes(Species,seconds(Mean_Time))) + geom_bar()  + theme_bw() + theme(axis.text.x=element_text(angle=-90)) + ylab("Average Seconds Feeding")

ggplot(dat,aes(x=seconds(Time_Feeder_Obs))) + geom_histogram()  + theme_bw()
ggsave(paste(gitpath,"Figures/Feedingtime.svg",sep=""),dpi=300,height=8,width=11)

```

Appendix A
============

Sampling information for each elevation

```{r,results='asis'}
print(xtable(vid_totals_date),type="html")
```
Table 1. Number of days of video analyzed for each elevation. For each elevation two sets of high and low value feeders were placed (O or R). The number of videos for each feeder that have been scored for hummingbird selectivity are shown for both the high and low value feeders. 


Selective matrix for all trials and species.

```{r,results='asis'}
print(xtable(selective.matrix),type="html")
```