---
title: "Appendix B"
author: "Ben Weinstein"
date: "Wednesday, January 21, 2015"
output:
  html_document:
    theme: spacelab
---
# Source Code

All code can be cloned from a github repository: 
https://github.com/bw4sz/Selectivity

To run code, just change the variable 'gitpath' to where you cloned the repo.

## Additional data files can be downloaded here:

Competition_Feeders:
https://www.dropbox.com/s/xss1hxz0wuhahl1/CompetitionFeeders.csv?dl=0

Hummingbird Morphology:
https://www.dropbox.com/s/m9zp2hdevi1vg5v/HummingbirdMorphology.csv?dl=0

Flower Transects: https://www.dropbox.com/s/j1awf90dzfi28oh/CleanedHolgerTransect.csv?dl=0

Flower Visitation:https://www.dropbox.com/s/0hdxpdkja8ph6t5/FlowerVideoClean.csv?dl=0


```{r,echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
opts_chunk$set(message=FALSE,warning=FALSE,fig.width=13,fig.height=10,echo=TRUE,cache=TRUE,cache.path = 'Source_cache/', fig.path='figure/')

library(ggplot2)
library(chron)
library(reshape)
library(reshape2)
library(plotKML)
library(maptools)
library(foreach)
library(vegan)
library(plyr)
library(dplyr)
library(rms)
library(MuMIn)
library(ape)
library(taxize)
library(bipartite)
load("Selectivity.RData")

```

#Set environment and read in data

```{r}
#set gitpath
gitpath<-"C:/Users/Ben/Documents/Selectivity/"

#source functions
source(paste(gitpath,"functions.R",sep=""))

#Set working directory
droppath<-"C:/Users/Ben/Dropbox/"

##Load image if needed
load("Selectivity.RData")

#Read in feeder data
dat<-read.csv(paste(droppath,"Thesis//Maquipucuna_SantaLucia/Data2013/csv/CompetitionFeeders.csv",sep=""))

#Make all entries capital
dat$Sex<-toupper(dat$Sex)

#Bring in Hummingbird Morphology Dataset, comes from
hum.morph<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/HummingbirdMorphology.csv",sep=""),row.names=1)

dat<-dat[,1:12]

#flower transect data
#read in flower totals from FlowerTransects.R
fltransects<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Results/FlowerTransects/CleanedHolgerTransect.csv",sep=""),row.names=1)

vid<-read.csv(paste(droppath,"Thesis/Maquipucuna_SantaLucia/Data2013/csv/FlowerVideoClean.csv",sep=""),row.names=1)



vid$Year<-years(chron(as.character(vid$Date)))
vid$Month<-as.numeric(months(chron(as.character(vid$Date))))

#gps introduced some duplicates, just remove any doubles based on time date species, id
toremove<-duplicated(vid[,c("ID","Time","Hummingbird","Iplant_Double","Month","Year")])

vid<-droplevels(vid[!toremove,])
#Bring in Phylogenetic Data
trx<-read.nexus(paste(gitpath,"InputData\\ColombiaPhylogenyUM.tre",sep=""))
spnames<-read.table(paste(gitpath,"InputData\\SpNameTree.txt",sep="") , sep = "\t", header = TRUE)

#Replace tip.label with Spnames#
#replace the tiplabels with periods, which is the biomod default
trx$tip.label<-gsub("_"," ",as.character(spnames$SpName))
ctrx<-cophenetic(trx)
```

#Data Check and Exploratory Analysis

```{r}
#Which dates need to be matched?
vid_totals_date<-aggregate(dat$Video,list(dat$Elevation,dat$Treatment,dat$Date,dat$Replicate),function(x) nlevels(droplevels(x)))
vid_totals_date<-dcast(vid_totals_date,Group.1 + Group.3 + Group.4~Group.2)

colnames(vid_totals_date)<-c("Elevation","Date","Replicate(O_R)","High Feeder","Low Feeder")

###General stats on sampling
#Number of days
paste("Number of days",nrow(vid_totals_date),sep=": ")

#no unknown species
dat<-dat[!dat$Species %in% "UKWN",]

#One misspelling of green crowned brilliant
levels(dat$Species)[levels(dat$Species) %in% "Green-crowned Brillaint"]<-"Green-crowned Brilliant"

#Create time columns
dat$Time.End<-chron::times(dat$Time.End)
dat$Time.Begin<-chron::times(dat$Time.Begin)

#Find time difference 
dat$Time_Feeder_Obs<-dat$Time.End - dat$Time.Begin

#Get any rownumbers that are negative, these need to be fixed. 
#dat[which(dat$Time_Feeder_Obs < 0),]

#average visits per hour
S_H<-table(hours(dat$Time.Begin),dat$Species)
```

##Species Presence over Time

```{r}
#Create overall date stamp
dat$Time_Stamp<-as.POSIXct(chron(dates=as.character(dat$Date),dat$Time.Begin))

#Time and species occurence, facetted by elevation
 ggplot(dat,aes(x=strptime(dat$Time.Begin,"%H:%M"),fill=Species)) + geom_histogram(position="dodge") + facet_wrap(~Elevation) + xlab("Time")
# ggsave("Thesis//Maquipucuna_SantaLucia/Results/TimeofDayElevation.svg",height=11,width=8,dpi=300)
# 
# #Overall Month_Day and Elevation
# ggplot(dat,aes(y=factor(Elevation),x=dat$Time_Stamp,col=Species)) + geom_point(size=3) + scale_x_datetime() + facet_wrap(~Species)
# ggsave("Thesis//Maquipucuna_SantaLucia/Results/DateElevation.svg",height=11,width=8,dpi=300)

#Species richness and identity at each elevation
sp_matrixHL<-(table(dat$Species,dat$Elevation,dat$Treatment) >= 1) * 1

#View species at each elevation and treatment
m.sp_m<-melt(sp_matrixHL)
colnames(m.sp_m)<-c("Species","Elevation","Treatment","Presence")

#turn 0's to NA's just for plotting
m.sp_m[m.sp_m$Presence==0,"Presence"]<-NA

#richness across feeders
rangefeed<-ggplot(m.sp_m,aes(y=Species,x=factor(Elevation),fill=as.factor(Presence)))+ geom_tile() + theme_bw() + scale_fill_discrete(na.value="white")
rangefeed + labs(fill="Present",x="Elevation")

#Total Time per species per elevation
Total_Time_Species<-aggregate(dat$Time_Feeder_Obs,by=list(dat$Species,dat$Elevation),sum,na.rm=TRUE) 
colnames(Total_Time_Species)<-c("Species","Elevation","TotalTime")
Total_Time_Species$Time<-minutes(Total_Time_Species$TotalTime)+seconds(Total_Time_Species$TotalTime)/60

p<-ggplot(Total_Time_Species,aes(y=Species,as.factor(Elevation),fill=Time)) + geom_tile() + theme_bw() + ylab("Minutes on Feeders") + theme(axis.text.x=element_text(angle=-90,vjust=-.1)) + scale_fill_continuous(low="gray90",high="gray10")
p+labs(x="Elevation",y="Species",fill="Minutes Feeding")

#mean time feeding bout
Mean_Time_Species<-aggregate(dat$Time_Feeder_Obs,by=list(dat$Species),mean,na.rm=TRUE) 
colnames(Mean_Time_Species)<-c("Species","Mean_Time")
```

```{r}
#Create a species list for each video
####Match each trial together, trials are done on the same day at the same elevation

#Split data into a list, with each compenent being one trial pair
Trials<-split(dat, list(dat$Elevation,dat$Date,dat$Replicate),drop=TRUE)

#Get number of levels per trial
levels.trial<-lapply(Trials,function(x) nlevels(factor(x$Treatment)))

#Only use trials that have a high and low, ie levels=2
complete.trial<- Trials[levels.trial ==2]

#time since feeding events
complete.trials<-lapply(1:length(complete.trial),function(g){
  #within feeding events
  tr<-complete.trial[[g]]
  
  order.x<-tr[order(tr$Time.Begin),]
  order.x$TimeSince<-NULL
  
  for(x in 2:nrow(order.x)){
    if(!order.x[x,"Species"] == order.x[x-1,"Species"]){
      ti<-order.x[x,]$Time.Begin - order.x[x-1,]$Time.End 
      order.x[x,"TimeSince"]<-minutes(ti)+seconds(ti)/60
    } else (order.x[x,"TimeSince"]<-NA)
  }
  
  order.x$Mass_diff<-NULL
  for(x in 2:nrow(order.x)){
    a<-order.x[x,"Species"]
    b<-order.x[x-1,"Species"]
    
    massA<-hum.morph[hum.morph$English %in% a, "Mass"]
    massB<-hum.morph[hum.morph$English %in% b, "Mass"]
    
    massDiff<-massB - massA
    if(length(massDiff) == 0){order.x[x,"Mass_diff"]<-NA} else{
      order.x[x,"Mass_diff"]<-massDiff
    }
  }
  return(order.x)
})

order.trials<-rbind.fill(complete.trials)
```


###Within trial metrics per species
```{r}
Tdata<-lapply(complete.trial,function(x){
  
  #caluclate selectivity
  a<-selective(x)
  
  #calculate birds per hour
  b<-bph(x)
  
  #Average time between feeding
  k<-time_feed(x)
  
  #average duration of feeding bout
  d<-avgF(x)
  
  #total visits
  tss<-data.frame(table(droplevels(x$Species)))
  colnames(tss)<-c("Species","N")
  
  #visits of other species
  visitsOthers<-sapply(levels(factor(x$Species)),function(g){
    sum(tss[!tss$Species %in% g,"N"])
  })
  
  visitsOthers<-data.frame(Species=names(visitsOthers),visitsOthers)
  

  #time since last species visit. 
    
  #merge data
  dat.trials<-merge(merge(merge(merge(merge(a,b),d),k),tss),visitsOthers)
  Elevation=unique(x$Elevation)
  Date=unique(x$Date)
  Trichness<-length(a[minutes(chron::times(a$Time_High + a$Time_Low)) > 1,]$Species)
  Replicate=unique(x$Replicate)
  
  #Total visits
  Tvisits<-nrow(x)
  
  out<-data.frame(dat.trials,Elevation,Date,Replicate,Richness=Trichness,Tvisits)
  return(out)})

#Bind dataset to a dataframe
selective.matrix<-rbind.fill(Tdata)
```

## Species visits per hour
```{r}
ggplot(selective.matrix,aes(x=Species,y=bph,fill=factor(Elevation))) + geom_boxplot()  + theme_bw() + facet_wrap(~Elevation,scales="free") + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + labs(y="Visits per hour",fill="Elevation")

#Create times argument
selective.matrix$Time_High<-chron::times(selective.matrix$Time_High)
selective.matrix$Time_Low<-chron::times(selective.matrix$Time_Low)
selective.matrix$Total_Time<-selective.matrix$Time_High + selective.matrix$Time_Low

#add total minutes feeding as a weight
selective.matrix$Minutes_High<-minutes(selective.matrix$Time_High) + seconds(selective.matrix$Time_High)/60
selective.matrix$Minutes_Low<-minutes(selective.matrix$Time_Low) + seconds(selective.matrix$Time_Low)/60
selective.matrix$Minutes_Total<-selective.matrix$Minutes_Low+selective.matrix$Minutes_High

#Add month column
selective.matrix$MonthA<-format(as.POSIXct(selective.matrix$Date,format="%m/%d/%Y"),"%b")
```

#Optimal foraging

Theory says that individuals should occupy patches at a rate equal to their quality
```{r}
sH<-sum(selective.matrix$Minutes_High)
sL<-sum(selective.matrix$Minutes_Low)

#Optimal foraging says its should be three
paste("Species feed on the high value", round(sH/sL,2), "times than the low value")

#set minimal number of datapoints. 
keep<-names(which(table(selective.matrix$Species) > 4))
```

###Total visits does not effect selectivity, just amount of total feeding

```{r}
ggplot(selective.matrix,aes(x=N,y=Minutes_High)) + geom_point() + stat_smooth(method="lm")
ggplot(selective.matrix,aes(x=N,y=Minutes_Low)) + geom_point() + stat_smooth(method="lm")
ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=N,y=Selectivity)) + geom_point() + ylim(0,1) + stat_smooth(method="lm") + facet_wrap(~Species,scales="free_x")
```

##Hummingbird Morphology and Selectivity

```{r}
#PCA of trait space
# Hum Standardized variables, what to do about NA's?
#Standard the matrix to correct for different units by subtracting the means and dividing by sd
zscore <- apply(hum.morph[,c("Bill","Mass","Tarsus_Length","Nail_Length","Wing_Loading")], 2, function(x){
  y<-(x - mean(x,na.rm=TRUE))/sd(x,na.rm=TRUE)
  return(y)
})
rownames(zscore)<-hum.morph$English

#Need to figure out what to do about Na's, we could use closely related species?
trait_pc<-prcomp(zscore)

#View Biplot of PC Space
biplot(trait_pc)

#distance in PCA space
d.pca<-as.matrix(dist(trait_pc$x))

#bind loadings 1 and 2 to dataframe
hum_load<-trait_pc$x[,c("PC1","PC2")]
rownames(hum_load)<-rownames(zscore)

#merge data
hum.morph<-merge(hum.morph,hum_load,by.x="English",by.y="row.names")

#merge PCA and original data with selectivity
selective.matrix<-merge(selective.matrix,hum.morph[,!colnames(hum.morph) %in% "Species"],by.x="Species",by.y="English")

#Weighted average of selectivity
ws<-sapply(split(selective.matrix,selective.matrix$Species),function(x){
  weighted.mean(x$Selectivity,x$Total_Time)
})

selective.matrix<-merge(selective.matrix,data.frame(weighted.selectivity=ws),by.x="Species",by.y="row.names")

#average metrics across species
avgStat<-aggregate(selective.matrix[,c("bph")],by=list(selective.matrix$Species),mean,na.rm=TRUE)

feeder<-avgStat[,-c(1)]

colnames(avgStat)[1]<-"Species"

#merge with weighted selectivity
avgStat<-merge(ws,avgStat,by.x="row.names",by.y="Species")
colnames(avgStat)<-c("Species","Weighted.Selectivity","Visits.per.hour")
rownames(avgStat)<-avgStat$Species

#take out any NA's
avgStat<-avgStat[complete.cases(avgStat),]

#take out species without enough data
avgStat<-avgStat[avgStat$Species %in% keep,]

pcaStat<-prcomp(avgStat[,-c(1)],scale=TRUE)
biplot(pcaStat)

#Facet by elevation
ggplot(selective.matrix,aes(x=Species,Selectivity,fill=Mass)) + geom_boxplot() + theme_bw() + theme(axis.text.x=element_text(angle=-90,vjust=-.1)) + facet_wrap(~Elevation,nrow=2,scale="free_x") + scale_fill_continuous(low="white",high="black") + labs(fill= "Mass (g)") 
#Facet by elevation, color by replicate
ggplot(selective.matrix,aes(x=Species,Selectivity,fill=Mass)) + geom_boxplot() + theme_bw()  + facet_wrap(~Elevation,scale="free_x",ncol=4) + scale_fill_gradient(low="blue",high="red")+ theme(axis.text.x=element_text(angle=65,hjust=1))
ggsave("Figures/Mass_SelectivityBoxplots.jpeg",dpi=300,height=8,width=11)
```

##Trait PCA and Selectivity
```{r,eval=TRUE}
traits<-hum.morph[,c("Bill","Mass","Wing_Loading","Nail_Length","Foot_Extension")]
rownames(traits)<-hum.morph$English

#aggregate by species
trait.f<-traits[rownames(traits) %in% avgStat$Species,]
```

```{r,eval=FALSE}
wss<-aggregate(selective.matrix$weighted.selectivity,by=list(selective.matrix$Species,selective.matrix$PC1,selective.matrix$PC2),mean)
colnames(wss)<-c("Species","PC1","PC2","weighted.selectivity")

biplot(prcomp(trait.f[rownames(trait.f) %in% keep,],scale=TRUE))
biplot(prcomp(trait.f[rownames(trait.f) %in% keep,],scale=TRUE))

```

##Selectivity as a function of difference in body size

```{r}
#Species list for each trial.
sp.lists<-sapply(complete.trial,function(x){
  levels(droplevels(x$Species))
})

#get body size lists for each trial
mass.lists<-lapply(sp.lists,function(x){
  hum.morph[hum.morph$English %in% x, "Mass"]
})


#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$MassD<-sapply(1:nrow(selective.matrix),function(y){
  
  x<-selective.matrix[y,]
  
  #find the index in the list
  index<-paste(paste(x["Elevation"],x[["Date"]],sep="."),x[["Replicate"]],sep=".")
  
  mass_T<-mass.lists[names(mass.lists) %in% index]
  
  
  #weighted mass difference, the mass to all species
  #multiple total time feeding by mass
  
  #get trial, with all species not included in the focal y row
  trialT<-selective.matrix[selective.matrix$Elevation %in% x[["Elevation"]]& selective.matrix$Date %in% x[["Date"]] & selective.matrix$Replicate %in% x[["Replicate"]] & !selective.matrix$Species==x[["Species"]] ,]  
  
  #create the mass "environment" ie mean mass difference between the focal species and all competitors, weighted by time feeding.
  #first get mass difference by each species and their time
  diff<-x$Mass - trialT$Mass
  
  diff<-diff[is.finite(diff)]
  #mean weighted differences
  weight.diff<-weighted.mean(diff,trialT$Minutes_Total,na.rm=TRUE)
  
  return(weight.diff)})

massplot<-ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=MassD,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(group=1))
massplot + facet_wrap(~Species,scale="free_x") 
```


##Selectivity as a function of phylogenetic relatedness

```{r}
#Species list for each trial.
sp.lists<-sapply(complete.trial,function(x){
  levels(droplevels(x$Species))
})


#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$PhyloD<-sapply(1:nrow(selective.matrix),function(y){
  x<-selective.matrix[y,]
  
  #find the index in the list
  index<-paste(paste(x["Elevation"],x[["Date"]],sep="."),x[["Replicate"]],sep=".")
  
  mass_T<-sp.lists[names(sp.lists) %in% index]
    
  #weighted phylo difference
  
  #get trial, with all species not included in the focal y row
  trialT<-selective.matrix[selective.matrix$Elevation %in% x[["Elevation"]]& selective.matrix$Date %in% x[["Date"]] & selective.matrix$Replicate %in% x[["Replicate"]] & !selective.matrix$Species==x[["Species"]] ,]  
  
  #create the mass "environment" ie mean mass difference between the focal species and all competitors, weighted by time feeding.
  #first get mass difference by each species and their time
  diff<-ctrx[rownames(ctrx) %in% x$double, colnames(ctrx) %in% trialT$double]
  
  if(length(diff)==0){return(NA)}
  
  if(length(diff)==1){
    return(diff)}
  else{
  #mean weighted differences
  weight.diff<-weighted.mean(diff,trialT[trialT$double %in% names(diff),"Minutes_Total"],na.rm=TRUE)
  return(weight.diff)}
  })

phyloplot<-ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=PhyloD,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(group=1))
phyloplot + facet_wrap(~Species,scale="free_x") 
```

##Selectivity and moving window of mass difference

```{r}
#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$MassD<-sapply(1:nrow(selective.matrix),function(y){
  
  x<-selective.matrix[y,]
  
  #find the index in the list
  index<-paste(paste(x["Elevation"],x[["Date"]],sep="."),x[["Replicate"]],sep=".")
  
  mass_T<-mass.lists[names(mass.lists) %in% index]
  
  
  #weighted mass difference, the mass to all species
  #multiple total time feeding by mass
  
  #get trial, with all species not included in the focal y row
  trialT<-selective.matrix[selective.matrix$Elevation %in% x[["Elevation"]]& selective.matrix$Date %in% x[["Date"]] & selective.matrix$Replicate %in% x[["Replicate"]] & !selective.matrix$Species==x[["Species"]] ,]  
  
  #create the mass "environment" ie mean mass difference between the focal species and all competitors, weighted by time feeding.
  #first get mass difference by each species and their time
  diff<-x$Mass - trialT$Mass
  
  diff<-diff[is.finite(diff)]
  #mean weighted differences
  weight.diff<-weighted.mean(diff,trialT$Minutes_Total,na.rm=TRUE)
  
  return(weight.diff)})

massplot<-ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=MassD,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1))
massplot + facet_wrap(~Species,scales="free_x",nrow=2) + theme_bw()
ggsave(paste(gitpath,"Figures/Mass_Selectivity.jpeg",sep=""),height=8,width=12,units="in",dpi=300)
```


##Selectivity as a function of difference in morph space

```{r}
#get body size lists for each trial
PC1.lists<-lapply(sp.lists,function(x){
  hum.morph[hum.morph$English %in% x, "PC1"]
})

#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$PC1D<-sapply(1:nrow(selective.matrix),function(y){
  
  x<-selective.matrix[y,]
  
  #find the index in the list
  index<-paste(paste(x["Elevation"],x[["Date"]],sep="."),x[["Replicate"]],sep=".")
  
  mass_T<-PC1.lists[names(PC1.lists) %in% index]
  
  #weighted mass difference, the mass to all species
  #multiple total time feeding by mass
  
  #get trial, with all species not included in the focal y row
  trialT<-selective.matrix[selective.matrix$Elevation %in% x[["Elevation"]]& selective.matrix$Date %in% x[["Date"]] & selective.matrix$Replicate %in% x[["Replicate"]] & !selective.matrix$Species==x[["Species"]] ,]  
  
  #create the morph "environment" ie mean mass difference between the focal species and all competitors, weighted by time feeding.
  #first get mass difference by each species and their time
  diff<-x$PC1 - trialT$PC1
  
  diff<-diff[is.finite(diff)]
  #mean weighted differences
  weight.diff<-weighted.mean(diff,trialT$Minutes_Total,na.rm=TRUE)
  
  return(weight.diff)})

PC1D<-ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=PC1D,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1))
PC1D + facet_wrap(~Species,scales="free_x")


#################Repeat for PC2######################
#####################################################
#Selectivity as a function of difference in body size
#####################################################

#get body size lists for each trial
PC2.lists<-lapply(sp.lists,function(x){
  hum.morph[hum.morph$English %in% x, "PC2"]
})


#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$PC2D<-sapply(1:nrow(selective.matrix),function(y){
  
  x<-selective.matrix[y,]
  
  #find the index in the list
  index<-paste(paste(x["Elevation"],x[["Date"]],sep="."),x[["Replicate"]],sep=".")
  
  mass_T<-PC2.lists[names(PC2.lists) %in% index]
  
  #weighted mass difference, the mass to all species
  #multiple total time feeding by mass
  
  #get trial, with all species not included in the focal y row
  trialT<-selective.matrix[selective.matrix$Elevation %in% x[["Elevation"]]& selective.matrix$Date %in% x[["Date"]] & selective.matrix$Replicate %in% x[["Replicate"]] & !selective.matrix$Species==x[["Species"]] ,]  
  
  #create the mass "environment" ie mean mass difference between the focal species and all competitors, weighted by time feeding.
  #first get mass difference by each species and their time
  diff<-x$PC2 - trialT$PC2
  
  diff<-diff[is.finite(diff)]
  #mean weighted differences
  weight.diff<-weighted.mean(diff,trialT$Minutes_Total,na.rm=TRUE)
  
  return(weight.diff)})

PC2D<-ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=PC2D,y=Selectivity,size=Minutes_Total,label=Species,col=Species)) + geom_point() + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total,group=1))
PC2D + facet_wrap(~Species,scales="free_x")

#################Multivariate distance######################
#####################################################
#Selectivity as a function of difference in body size
#####################################################

#For each row in the selectivity matrix, get the difference to the largest species

#get the species with the max body size at the feeder during that trial
selective.matrix$MultD<-sapply(1:nrow(selective.matrix),function(y){
  
  x<-selective.matrix[y,]
  
  #weighted mass difference, the mass to all species
  #multiple total time feeding by mass
  
  #get trial, with all species not included in the focal y row
  trialT<-selective.matrix[selective.matrix$Elevation %in% x[["Elevation"]]& selective.matrix$Date %in% x[["Date"]] & selective.matrix$Replicate %in% x[["Replicate"]] & !selective.matrix$Species==x[["Species"]] ,]  
  
  #get mean multivariate distance
  PCdiff<-d.pca[as.character(x[["Species"]]),as.character(trialT[["Species"]])]
  
  PCdiff<-PCdiff[is.finite(PCdiff)]
  #mean weighted differences
  weight.diffW<-weighted.mean(PCdiff,trialT$Minutes_Total,na.rm=TRUE)
  weight.diff<-mean(PCdiff,na.rm=TRUE)
  
  return(weight.diffW)})

MultD<-ggplot(data=selective.matrix[selective.matrix$Species %in% keep,],aes(x=MultD,y=Selectivity,size=Minutes_Total,col=Species)) + stat_smooth(method="glm",family="binomial",aes(weight=Minutes_Total)) + geom_point()
MultD + facet_wrap(~Species,scales="free_x")
```

##Selectivity and Available Resource

```{r}
#just summer data
fltransects<-droplevels(fltransects[fltransects$month %in% c(6,7,8) & fltransects$Year %in% "2013",])

fltransects$GPS_ID<-as.numeric(as.character(fltransects$GPS_ID))

#just get videos from the summer months
vid<-vid[vid$Month %in% c(6,7,8) & vid$Year %in% 2013,]

#match with spatial infodim
gps<-readShapePoints(paste(droppath,"Thesis\\Maquipucuna_SantaLucia\\Data2013\\Shapefiles\\GPSshape.shp",sep=""))@data

gps<-droplevels(gps[gps$MonthID %in% c(6,7,8),])

gps$GPS_ID<-as.numeric(as.character(gps$GPS_ID))

#GPSID + month
gps$mergeID<-paste(gps$GPS_ID,gps$MonthID,sep="_")

#remove duplicated
gps_d<-gps[duplicated(gps,fromLast=TRUE),]

fltransects$mergeID<-paste(fltransects$GPS_ID,fltransects$month,sep="_")

fltransects<-merge(fltransects,gps_d,by="mergeID",all.x=TRUE)

#where elevation failed, take the midpoint of the transect
fltransects[is.na(fltransects$ele),"ele"]<-apply(fltransects[is.na(fltransects$ele),c("Elevation.Begin","Elevation.End")],1,mean)

#Split videos into species lists
vid.s<-split(vid,list(vid$Hummingbird))
selective.matrix[,c("fl_all", "fl_s")]<-foreach(g=1:nrow(selective.matrix),.combine=rbind) %do%{
  
  #grab row
  x<-selective.matrix[g,]
  
  #Get the hummingbird index
  flIndex<-vid.s[[as.character(x$Species)]]
  
  #Which species does the bird feed on
  fl.sp<-flIndex$Iplant_Double
  
  #dates within two weeks either side. 
  dateL<-abs(difftime(as.POSIXlt(x[["Date"]],format="%m/%d/%Y"),  as.POSIXlt(fltransects$Date,format="%m/%d/%Y"),units="weeks")) <= 2.5
  
  #and within 200m of the elevation
  flower.month<-fltransects[dateL & (fltransects$ele > x[["Elevation"]]-100 &fltransects$ele < x[["Elevation"]]+ 100) ,]
  
  if(nrow(flower.month)==0) return(NA)
  
  mean.fl_all<-mean(aggregate(flower.month$Total_Flowers,list(flower.month$Transect.ID),sum,na.rm=TRUE)$x)
    
  flower_sp<-flower.month[flower.month$Iplant_Double %in% fl.sp,]
  
  if(nrow(flower_sp)==0) return(NA)
  
  mean.fl<-mean(aggregate(flower_sp$Total_Flowers,list(flower_sp$Transect.ID),sum,na.rm=TRUE)$x)
  
  return(c(mean.fl_all,mean.fl))}

flowerplot<-melt(id.var=c("Species","Selectivity","Minutes_Total","bph"),measure.vars=c("fl_s","fl_all"),data=selective.matrix)


resourceplotS<-ggplot(flowerplot[flowerplot$Species %in% keep,],aes(x=value,linetype=variable,y=Selectivity,col=variable,shape=variable,size=Minutes_Total)) + geom_point() +  stat_smooth(show_guide=F,method="glm",family="binomial",aes(weight=Minutes_Total,group=variable,col=variable))

resourceplotS+ facet_wrap(~Species,scales="free_x",nrow=3) + labs(x="Available Resources") + theme_gray() + scale_color_manual(values=c("black","grey40"),"Resource Model",labels=c("Preferred Resources","All Resources")) + scale_shape_manual("Resource Model",labels=c("Preferred Resources","All Resources"),values=c(16,17)) + scale_linetype_discrete(guide="none") + labs(size="Minutes") + theme_bw()

ggsave("Figures/Resource_Selectivity.jpeg",dpi=300,height=7,width=10.5,units="in")
```

#Model Fitting

```{r}
selective.matrix$MassD[!is.finite(selective.matrix$MassD)]<-0
selective.matrix$MultD[!is.finite(selective.matrix$MultD)]<-0

mod_data<-selective.matrix[selective.matrix$Species %in% keep & complete.cases(selective.matrix[,c("Species","Mass","MassD","PhyloD","MultD","fl_all","fl_s")]), ]

missing<-selective.matrix[!complete.cases(selective.matrix[,c("Species","Mass","MassD","MultD","PhyloD", "fl_all","fl_s")]), ]

##Compute model fits

#Create response column, needs to be formatted, cbind(successes,failures)
response<-cbind((minutes(mod_data$Time_High)+seconds(mod_data$Time_High)/60),(minutes(mod_data$Time_Low)+seconds(mod_data$Time_Low)/60))

#Intercept only
m_intercept<-glm(data=mod_data,response~1,family="binomial",na.action="na.fail")

#By Species
m_species<-glm(data=mod_data,response~Species,family="binomial",na.action="na.fail")

#By Mass
m_mass<-glm(data=mod_data,response~Mass*Species,family="binomial",na.action="na.fail")

#By Difference in Mass
m_massD<-glm(data=mod_data,response~MassD*Species,family="binomial",na.action="na.fail")

#By multivariate distance
m_multD<-glm(data=mod_data,response~MultD*Species,family="binomial",na.action="na.fail")

#By phylogenetic distance
m_phyloD<-glm(data=mod_data,response~PhyloD*Species,family="binomial",na.action="na.fail")

#By available resources, all
m_Allresources<-glm(data=mod_data,response~fl_all*Species,family="binomial",na.action="na.fail")

#By available resources, by species
m_Speciesresources<-glm(data=mod_data,response~fl_s*Species,family="binomial",na.action="na.fail")

mods<-list(m_intercept,m_species,m_mass,m_massD,m_phyloD,m_multD,m_Allresources,m_Speciesresources)
```

##Model Criticism

###AIC
```{r}
missing<-selective.matrix[!selective.matrix$Species %in% keep & complete.cases(selective.matrix[,c("Species","Mass","MassD","MultD","fl_all")]), ]

aictable<-AIC(m_intercept,m_species,m_mass,m_massD,m_multD,m_Allresources,m_phyloD,m_Speciesresources)
aictable$Delta<-aictable$AIC-min(aictable$AIC)
aictable[order(aictable$AIC),]

capture.output(aictable,file="Figures/AIC.xls")

#write table to file
#capture.output(at,file="Figures/Anova.xls")
```

#Multimodel inference

```{r}
modelA<-model.avg(m_intercept,m_species,m_mass,m_massD,m_multD,m_Allresources,m_Speciesresources,m_phyloD)

#get N
mods<-list(m_intercept,m_species,m_mass,m_massD,m_multD,m_Allresources,m_Speciesresources,m_phyloD)

#Get psuedoR2

R2<-sapply(mods,function(x){
  1-x$deviance/x$null.deviance
})

sumAIC<-round(modelA$msTable,2)
print(sumAIC)
modelA

###############Needs to be hand ordered if rerun, no match yet.
sumAIC$PsuedoR2<-round(unlist(R2),2)[c(7,6,8,5,2,3,4,1)]

sumAIC
capture.output(sumAIC,file="Figures/AIC.xls")

#anova(m_Speciesresources,m_MorphResources,test="Chisq")
```


#Mantel test of traits and resources

```{r}
# #format flower matrix
flmatrix<-as.data.frame.array(table(vid$Hummingbird,vid$Iplant_Double))

# remove all 0 rows
toremove<-names(which(apply(flmatrix,2,sum)==0))

flmatrix<-flmatrix[,!colnames(flmatrix) %in% toremove]

#match the matrices
###Morphology and Phylogeny in predicting trait matrix
flower<-flmatrix[rownames(flmatrix) %in% rownames(trait.f),]

# remove all 0 rows
toremove<-names(which(apply(flower,2,sum)==0))

flower<-flower[,!colnames(flower) %in% toremove][,-1]

mantel(dist(trait.f[,-3]),dist(flower))

#phylogenetic distance
#species in the flower matrix
sp_keep<-data.frame(E=rownames(flower),S=unique(selective.matrix[selective.matrix$Species %in% rownames(flower),"double"]))

phy_dist<-ctrx[rownames(ctrx) %in% sp_keep$S,colnames(ctrx) %in% sp_keep$S]

#reorder to match flower matrix
ord<-as.character(sp_keep$S)
re_ord<-as.numeric(sapply(ord,function(x) which(x == rownames(phy_dist))))
phy_ord<-phy_dist[re_ord,re_ord]

mantel(phy_ord,dist(flower))
```

#Network of Hummingbird Plant Interactions

```{r}
#plot in the same order
orderflowers<-names(sort(apply(flower,2,sum),decreasing=FALSE))
orderbirds<-names(sort(apply(flower,1,sum),decreasing=TRUE))

a<-melt(as.matrix(flower))

colnames(a)<-c("Birds","Flowers","value")

a$Flowers<-factor(a$Flowers,levels=orderflowers)
a$Birds<-factor(a$Birds,levels=orderbirds)

ggplot(a[a$value>0,],aes(x=Birds,y=Flowers,fill=value)) + geom_tile()+ theme_bw() + scale_fill_continuous(low="grey90",high="black") + theme(axis.text.x = element_text(angle = 60, hjust = 1)) + labs(fill="# of Visits")
ggsave("Figures/MatrixPlot.jpeg",dpi=600,height=7,width=6)
```

#Visitation rates
```{r}
head(selective.matrix)

gvisit<-melt(selective.matrix, id.vars=c("Species","Elevation","Date","Replicate","Selectivity","Tvisits","bph","Minutes_Total","fl_s","fl_all"),measure.vars=c("Minutes_High","Minutes_Low"))

ggplot(gvisit[gvisit$Species %in% keep,],aes(x=fl_s,y=value,col=variable)) + geom_point(size=2) + facet_wrap(~Species,scale="free") + geom_smooth(method="glm",family="poisson",alpha=.1) + theme_bw() + labs(x="Abundance of Preferred Resources",y="Minutes on Feeder",col="Feeder") + scale_color_manual(labels=c("High Value","Low Value"),values=c("Black","grey60"))

#Total minutes
gvisit<-melt(selective.matrix, id.vars=c("Species","Elevation","Date","Replicate","Selectivity","Tvisits","bph","Minutes_Total","Minutes_High","Minutes_Low"),measure.vars=c("fl_all","fl_s"))

ggplot(selective.matrix[selective.matrix$Species %in% keep,],aes(x=fl_s,y=bph)) + geom_point() + facet_wrap(~Species,scale="free") + geom_smooth(method="glm",family="poisson",col="black") + theme_bw() + labs(x="Log Abundance of Preferred Resources",y="Birds per hour")
ggsave("Figures/bphresources.jpg",height=5,width=9,dpi=600)
```
#Analysis of Niche Overlap

```{r}
FH<-t(as.matrix(flower))
networkL<-networklevel(FH,"niche overlap")[[1]]
null_niche<-sapply(nullmodel(FH),function(x){
  networklevel(x,"niche overlap")[[1]]})

ggplot() + geom_histogram(data=data.frame(null_niche),aes(x=null_niche)) + theme_bw() + geom_vline(xintercept=networkL,linetype='dashed',col='grey50',size=1.5) + xlab("Niche Overlap") 
ggsave("Figures/NicheOverlap.jpg",units="in",width=5,height=3,dpi=600)

```


#Displacement analysis

How many seconds between feeding events?
We will do this for each row in each trial.

```{r}
sdat<-split(dat,list(dat$Date,dat$Elevation,dat$Replicate,dat$Treatment),drop = T)
datd<-rbind.fill(lapply(sdat,function(x){
  x<-x[order(x$Time_Stamp),]
  #make column to fill with time stampa difference
  x$DiffTime<-NA
  for(d in 1:(nrow(x)-1)){
  x$DiffTime[d+1]<-as.numeric(difftime(x$Time_Stamp[d+1],x$Time_Stamp[d],units = "secs"))
  }
  return(x)
}))

#for each displacement event get information and change

dspl<-which(datd$DiffTime<30)

out<-list()
for(i in 1:length(dspl)){
  x<-dspl[i]
out[[x]]<-data.frame(Date=datd[x,"Date"],Replicate=datd[x,"Replicate"],Treatment=datd[x,"Treatment"],Elevation=datd[x,"Elevation"],Holder=datd[x-1,"Species"],Intruder=datd[x,"Species"])  
}
displ.df<-rbind.fill(out)

displm<-melt(table(displ.df$Holder,displ.df$Intruder,displ.df$Treatment))

#take out intraspecific since we can't tell if its the same individual
colnames(displm)<-c("Holder","Intruder","Treatment","value")
displm<-displm[!displm$Holder==displm$Intruder,]
displm<-displm[displm$Treatment %in% c("H","L"),]

#create come meaningful order
#doesn't like the name value
colnames(displm)[4]<-"val"
hord<-displm %>% group_by(Holder) %>% summarise(s=sum(val)) %>% arrange(s) %>% select(Holder) %>% as.vector()

displm$Holder<-factor(displm$Holder,levels=hord$Holder)

Iord<-displm %>% group_by(Intruder) %>% dplyr::summarize(s=sum(val)) %>% arrange(desc(s)) %>% select(Intruder) %>% as.vector()
displm$Intruder<-factor(displm$Intruder,levels=Iord$Intruder)

#remove all non interactions
displm<-droplevels(displm[!displm$val==0,])

#plot
ggplot(displm,aes(x=Intruder,y=Holder,fill=val)) + geom_tile() + scale_fill_continuous(low="blue",high="red") + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + facet_wrap(~Treatment) + labs(fill="Displacements")

#as a function of relatedness and morphological similarity
#phylo
displm$Phydist<-NA
for(x in 1:nrow(displm)){
  #get scientific name
  h<-displm[x,"Holder"]
  hsp<-unique(selective.matrix[selective.matrix$Species %in% h,"Row.names"])
  
  I<-displm[x,"Intruder"]
  Isp<-unique(selective.matrix[selective.matrix$Species %in% I,"Row.names"])
  #if we don't have records, then skip
  if(length(hsp)==0|length(Isp)==0){next}
  displm$Phydist[x]<-ctrx[hsp,Isp]
  displm$Morphdist[x]<-d.pca[displm[x,"Holder"],displm[x,"Intruder"]]
}

displm<-melt(displm,measure.vars = c("Morphdist","Phydist"))
colnames(displm)[6]<-c("distance")
ggplot(displm,aes(x=distance,y=val,col=variable)) + geom_point(size=3) + geom_smooth(method="glm",family="poisson") + facet_grid(Treatment~variable,scales="free") 
```

```{r}
#save data
save.image("Selectivity.RData")
```
